Improving the Image Acquisition Rate of an Atomic Force Microscope Through Compressive Sampling

### 2019-05-28 latex/afm-cs-main.tex
Symlink my bib file into repo.
Symlink doesnt work.
Makefile for block-diagrams.

\documentclass[twocolumn,oneside]{IEEEtran/IEEEtran}
% \documentclass[defaultstyle,11pt]{IEEETran}
% Lucy says to look at:
% float_page_fraction
% tex_page_fraction
% https://tex.stackexchange.com/questions/35125/how-to-use-the-placement-options-t-h-with-figures#35130
% \renewcommand\floatpagefraction{0.5} %% default: 0.5
% \renewcommand\topfraction{0.7} % default: 0.7
% \renewcommand*\textfraction{.05}
\usepackage{showframe}
\usepackage{amssymb}		% to get all AMS symbols
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{graphicx}		% to insert figures
\usepackage{hyperref}		% PDF hyperreferences??
\usepackage{multirow}
\usepackage{mathtools}
\usepackage{cite}
\usepackage{listings}
\usepackage{booktabs}
% \usepackage{amsthm}
\usepackage[percent]{overpic}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{xspace}
\usepackage{enumitem}
\usepackage{catchfilebetweentags} % for data input
\usepackage[inkscapelatex=false, inkscapepath=svgsubpath]{svg}
\usepackage{url}
\usepackage{mdframed}
\usepackage{amsmath, dsfont, bbold}
% \usepackage{algpseudocode,algorithm,algorithmicx}
\usepackage[ruled,linesnumbered]{algorithm2e}
\usepackage{placeins}
\usepackage{todonotes}

@@

% =========================================================================
\section{Experimental setup} \label{sec:experimentalSetup}
% =========================================================================
Our experimental setup, illustrated in Fig.~\ref{fig:exp_setup_anderssonLab},
consists of an Agilent 5400 AFM  retrofitted with an nPoint NPXY100A
piezoelectric stage, a cRIO-9082 embedded controller (National
Instruments), and a standard desktop computer.

@@

All control logic is programmed using LabVIEW 2019 and compiled to a Xilinx
Spartan-6 LX150 Field Programmable Gate Array (FPGA) inside the cRIO-9082. The
cRIO includes a 16-Bit, 100 kHz NI-9215 analog-to-digital input module and a
16-Bit, 100 kHz NI-9263 digital-to-analog output module. All control loops are
implemented using a 25 kHz sampling rate.

% =========================================================================
\section{Control}\label{sec:control}
The control structure for each axis is different. For the $X$-axis, we use a
state space controller. It is described fully in \cite{braker_afmmpc_2019} and
achieves a bandwidth of about 150 Hz, which is over 1/3 of the first $X$-axis
resonance at 350 Hz.

@@

\begin{figure*}[t]
  \centering
  \includegraphics[width=0.75\textwidth]{figures/AFM_loop_Dinv_antiwup-crop.pdf}
  \caption{Block diagram of the $Z$-axis control loop. The anti-windup is active
    for decision index $i\in\{0,1,5\}$, which correspond to states such that
    $r_Z=\rzup$.}
  \label{fig:afm_bd_dinv}
\end{figure*}

@@

\begin{table}\label{tab:cs_tasks}
  \centering
  \caption{Summary of the tasks for $\mu$-path scanning.}
  \begin{tabular}{lccc}
    state & $X$,$Y$ & $Z$ & next-state\\
    \toprule
    (1) $XY$-move & setpoint & $r_Z = \rzup$ & (2)\\
    (2) tip-engage & setpoint & $r_Z = \rzs$ & (3)\\
    (3) pre-scan & ramp & $r_Z = \rzs$ & (4)\\
    (4) $\mu$-path scan & ramp & $r_Z = \rzs$ & (5)\\
    (5) tip up & setpoint & $r_Z = \rzup$ & (1)\\
  \end{tabular}
\end{table}
Two cycles of this sequential process are illustrated by the time series in
Fig.~\ref{fig:mupathsignaltrajectory}. The following subsections describe in
more detail several features of the CS scanning process.

@@

\subsection{The pre-scan}
\begin{figure*}
  \begin{subfigure}{.33\textwidth}
    \includesvg[width=1\textwidth]{figures/zbounce_uz_example.svg}
    \caption{ }
    \label{fig: }
  \end{subfigure}
  \begin{subfigure}{.33\textwidth}
    \includesvg[width=1\textwidth]{figures/prescan_uz_example.svg}
    \caption{ }
    \label{fig:uz_prescan}
  \end{subfigure}
  \begin{subfigure}{.33\textwidth}
    \includesvg[width=1\textwidth]{figures/prescan_zbounce_PSD.svg}
    \caption{ }
    \label{fig: }
  \end{subfigure}
  \caption{(a) A ``$Z$-bounce'' experiment where the $XY$-stage is motionless
    and the cantilever is repeatedly engaged and disengaged. (b) One cycle of a
    tightly packed set of CS-cycles where the movement between each measurement
    location is 0. (c) PSDs from the portions of the trajectories which are
    colored blue and turquoise, averaged over many such cycles, demonstrating
    that the scan process itself reduces vibration in the $Z$-axis.}
  \label{fig:prescan_difference}
\end{figure*}
After the tip-descent, we begin scanning in a ``pre-scan'' phase (state (3)).
During this phase,

To
motivate, this consider Fig.~\ref{fig:prescan_difference}. The left panel
represents what we call a ``$Z$''-bounce experiment, where the $XY$-stage stays
motionless, but we repeatedly retract and re-engage the tip with specimen,
exactly as we would in a CS scan. The right panel represents a single cycle from
specialized CS pattern, where the $\mu$-paths are still 500 nm long, but the
paths are separated by 0 pixels.
In this case, the specimen is freshly cleaved
mica.
The goal here is to eliminate (i) any oscillations which arise from the
$XY$-move (which, as we will show in Section ~\ref{sec:need_mimo}, can be
significant) and (ii) to have a perfectly flat surface to scan over so that
there are as few disturbances as possible entering the control loop. Finally,
the left panel of Fig.~\ref{fig:prescan_difference} shows the PSDs of the left
and right panels (but averaged over many similar cycles). Note that the PSD is
only computed from the combination of the blue and turquoise signals.

@@

\todo{Should probably remove this entire section?}

What kind of numbers are reasonable to expect from the SSIM and PSNR metrics? We
mentioned in Chapter~\ref{chap:cs_init} that comparison of experimental images
was difficult (particularly for our experimental setup) because, e.g., Agilent
piezo tube drifts which gives a varying offset between images. However, we did
not try to tease apart to what extent the poor SSIM and PSNR numbers we reported
were due to simply the difference in scanning speeds vs things like image
mis-alignment.

@@

In the first experiment, (Figs \ref{fig:rast_unaligned_0p5} and
\ref{fig:rast_aligned_0p5}) we took 6 raster scans of the same 5 by 5 micron
square of the CS-20NG sample grating. These scans were taken sequentially and
the NPXY100A remained in closed-loop throughout the process. Taking the first
scan as the master, we compute (without any alignment) the PSNR and SSIM
metrics. Fig.~\ref{fig:rast_unaligned_0p5} shows the error between the master
image and the remaining 6 images. Note that the best PSNR here is 12.13 and the
worst is 7.87. The best SSIM is 0.40 and the worst is 0.14.

Next, using the cross-correlation alignment technique discussed in
Section~\ref{sec:align}, we aligned the master image with a sub-slice of each of
the remaining 6 (the sub-slice removes 25 pixels from each side). Again, we
computed the PSNR and SSIM. These results are shown in
Fig.~\ref{fig:rast_aligned_0p5}, which again shows the errors between the
aligned master and the remaining 6 images. Although the situation is improved
compared to the un-aligned images, there is still noticable stretching,
particularly in the $Y$-direction. Notably, the with the alignment technique,
PSNR now ranges from SSIM 21.68 to 15.56 and the SSIM ranges from 0.67 to 0.50.

We repeat this experiment but this time use a 1~Hz scan.
Fig.~\ref{fig:rast_unaligned_1} shows the unaligned error images and
Fig.\ref{fig:rast_aligned_1} shows the aligned error images. Temporally, the
images are order left-to-right and top-to-bottom. Thus we can see, both visually
and via the metrics, that in Fig.~\ref{fig:rast_unaligned_1} the mis-alignment
grows over time (In the 0.5~Hz scans, it appears that after the third from, the
misalignment is so bad, the situation can't get any worse). In the aligned 1~Hz
images, our metrics are not only higher but also more consistent: and the SSIM
lies between 0.7 and 0.73 and the PSNR ranges from 22.51 to 24.99.

@@

For the raster scans, we discard data from the re-trace and divided the
remaining data into 512$\times$512 bins based on the \xc and \yc sensor
measurements. We then average the data in each bin to obtain the value of one
pixel. To remove the affects of piezo creep and sample tilt, we then de-trend
each individual line. De-trending each line introduces the artifacts we noted in
Section~\ref{sec:comp_raster}. To remove this effect, we select a column of
pixels on the left and right side of each image that does not cross any holes
(i.e., a vertical line through the flat area), register each scan line to a
common height along these columns. This step eases comparison with the CS scans
and was inspired by a feature in \texttt{SPIW} \cite{spiw}.

@@

%%%%%%%%% then the Bibliography, if any %%%%%%%%%
% \bibliographystyle{plain} % or "siam", or "alpha", etc.
\bibliographystyle{IEEEtran} % or "siam", or "alpha", etc.
% \nocite{*} % list all refs in database, cited or not
\bibliography{/home/arnold/bib_pdf/main_bibliography,Andersson_bibs,Yufans_bibs} % Bib database in "refs.bib"

>>>>

\documentclass[twocolumn,oneside]{IEEEtran/IEEEtran}
% \documentclass[defaultstyle,11pt]{IEEETran}
% Lucy says to look at:
% float_page_fraction
% tex_page_fraction
% https://tex.stackexchange.com/questions/35125/how-to-use-the-placement-options-t-h-with-figures#35130
% \renewcommand\floatpagefraction{0.5} %% default: 0.5
% \renewcommand\topfraction{0.7} % default: 0.7
% \renewcommand*\textfraction{.05}
\usepackage{showframe}
\usepackage{amssymb}		% to get all AMS symbols
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{graphicx}		% to insert figures
% \usepackage{hyperref}		% PDF hyperreferences??
\usepackage{multirow}
\usepackage{mathtools}
\usepackage{cite}
\usepackage{listings}
\usepackage{booktabs}
% \usepackage{amsthm}
\usepackage[percent]{overpic}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{xspace}
\usepackage{enumitem}
\usepackage{catchfilebetweentags} % for data input
\usepackage[inkscapelatex=false, inkscapepath=svgsubpath]{svg}
\usepackage{url}
\usepackage{mdframed}
\usepackage{amsmath, dsfont, bbold}
% \usepackage{algpseudocode,algorithm,algorithmicx}
\usepackage[ruled,linesnumbered]{algorithm2e}
\usepackage{placeins}
\usepackage{todonotes}

@@

% =========================================================================
\section{Experimental setup} \label{sec:experimentalSetup}
% =========================================================================
Our experimental setup, illustrated in Fig.~\ref{fig:exp_setup},
consists of an Agilent 5400 AFM  retrofitted with an nPoint NPXY100A
piezoelectric stage, a cRIO-9082 embedded controller (National
Instruments), and a standard desktop computer.

@@

All control logic is programmed using LabVIEW 2019 and compiled to a Xilinx
Spartan-6 LX150 Field Programmable Gate Array (FPGA) inside the cRIO-9082. The
cRIO includes a 16-Bit, 100 kHz NI-9215 analog-to-digital input module and a
16-Bit, 100 kHz NI-9263 digital-to-analog output module. All control loops are
implemented using a 25 kHz sampling rate.
\begin{figure}
  \includegraphics[width=1\columnwidth]{figures/exp_setup.pdf}
  \caption{Schematic diagram of the expermental AFM setup.}
  \label{fig:exp_setup}
\end{figure}
% =========================================================================
\section{Control}\label{sec:control}
The control structure for each axis is different. For the $X$-axis, we use a
state space controller. It is described fully in \cite{braker_afmmpc_2019} and
achieves a bandwidth of about 150 Hz, which is over 1/3 of the first $X$-axis
resonance at 350 Hz.

@@

\begin{figure*}[t]
  \centering
  \includegraphics[width=0.75\textwidth]{figures/AFM_loop_z.pdf}
  \caption{Block diagram of the $Z$-axis control loop. The anti-windup is active
    for decision index $i\in\{0,1,5\}$, which correspond to states such that
    $r_Z=\rzup$.}
  \label{fig:afm_bd_dinv}
\end{figure*}

@@

\begin{table}
  \centering
  \caption{Summary of the tasks for $\mu$-path scanning.}
  \begin{tabular}{lccc}
    state & $X$,$Y$ & $Z$ & next-state\\
    \toprule
    (1) $XY$-move & setpoint & $r_Z = \rzup$ & (2)\\
    (2) tip-engage & setpoint & $r_Z = \rzs$ & (3)\\
    (3) pre-scan & ramp & $r_Z = \rzs$ & (4)\\
    (4) $\mu$-path scan & ramp & $r_Z = \rzs$ & (5)\\
    (5) tip up & setpoint & $r_Z = \rzup$ & (1)\\
  \end{tabular}
  \label{tab:cs_tasks}
\end{table}
Two cycles of this sequential process are illustrated by the time series in
Fig.~\ref{fig:mupathsignaltrajectory}. The following subsections describe in
more detail several features of the CS scanning process.

@@

\subsection{The pre-scan}
\begin{figure*}
  \begin{subfigure}{.33\textwidth}
    \includesvg[width=1\textwidth]{figures/zbounce_uz_example.svg}
    \caption{ }
    % \label{fig: }
  \end{subfigure}
  \begin{subfigure}{.33\textwidth}
    \includesvg[width=1\textwidth]{figures/prescan_uz_example.svg}
    \caption{ }
    \label{fig:uz_prescan}
  \end{subfigure}
  \begin{subfigure}{.33\textwidth}
    \includesvg[width=1\textwidth]{figures/prescan_zbounce_PSD.svg}
    \caption{ }
    % \label{fig: }
  \end{subfigure}
  \caption{(a) A ``$Z$-bounce'' experiment where the $XY$-stage is motionless
    and the cantilever is repeatedly engaged and disengaged. (b) One cycle of a
    tightly packed set of CS-cycles where the movement between each measurement
    location is 0. (c) PSDs from the portions of the trajectories which are
    colored blue and turquoise, averaged over many such cycles, demonstrating
    that the scan process itself reduces vibration in the $Z$-axis.}
  \label{fig:prescan_difference}
\end{figure*}
After the tip-descent, we begin scanning in a ``pre-scan'' phase (state (3)).
During this phase,

To motivate, this consider Fig.~\ref{fig:prescan_difference}. The left panel
represents what we call a ``$Z$''-bounce experiment, where the $XY$-stage stays
motionless, but we repeatedly retract and re-engage the tip with specimen,
exactly as we would in a CS scan. The right panel represents a single cycle from
specialized CS pattern, where the $\mu$-paths are still 500 nm long, but the
paths are separated by 0 pixels.
In this case, the specimen is freshly cleaved
mica.
The goal here is to eliminate (i) any oscillations which arise from the
$XY$-move and (ii) to have a perfectly flat surface to scan over so that
there are as few disturbances as possible entering the control loop. Finally,
the left panel of Fig.~\ref{fig:prescan_difference} shows the PSDs of the left
and right panels (but averaged over many similar cycles). Note that the PSD is
only computed from the combination of the blue and turquoise signals.

@@

\subsection{Limitations of SSIM and PSNR}\label{sec:limits}
\todo{Should probably remove this entire section?}

What kind of numbers are reasonable to expect from the SSIM and PSNR metrics?
To answer this question, we compare a sequence of 1~Hz raster scans.
In this experiment, (Fig. \ref{fig:rast_unaligned_1}) we took 6 raster scans of the same 5 by 5 micron
square of the CS-20NG sample grating. These scans were taken sequentially and
the NPXY100A remained in closed-loop throughout the process.
Fig.~\ref{fig:rast_unaligned_1} shows the error between the master
image and the remaining 6 images.
Temporally, the images are order left-to-right and top-to-bottom.
Thus we can see, both visually and via the metrics, that in Fig.~\ref{fig:rast_unaligned_1} the mis-alignment
grows over time.

Taking the first scan as the master, we compute the PSNR and SSIM
metrics. The best PSNR here is 16.57 and the worst is 9.02.
The best SSIM is 0.54 and the worst is 0.28.

@@

Next, using a cross-correlation alignment technique
% discussed in Section~\ref{sec:align},
we aligned the master image with a sub-slice of each of
the remaining 6 (the sub-slice removes 25 pixels from each side). Again, we
computed the PSNR and SSIM. These results are shown in
Fig.~\ref{fig:rast_aligned_1}, which again shows the errors between the
aligned master and the remaining 6 images. Although the situation is improved
compared to the un-aligned images, there is still noticable stretching,
particularly in the $Y$-direction. Notably, with the alignment technique,
the SSIM lies between 0.7 and 0.73 and the PSNR ranges from 22.51 to 24.99 and appears more stable in time.

@@

For the raster scans, we discard data from the re-trace and divided the
remaining data into 512$\times$512 bins based on the \xc and \yc sensor
measurements. We then average the data in each bin to obtain the value of one
pixel. To remove the affects of piezo creep and sample tilt, we then de-trend
each individual line. De-trending each line causes the flat portion between the holes in the horizontal direction to appear at a different height than the flat portion between holes in the vertical direction. 
To remove this effect, we select a column of
pixels on the left and right side of each image that does not cross any holes
(i.e., a vertical line through the flat area), register each scan line to a
common height along these columns. This step eases comparison with the CS scans
and was inspired by a feature in \texttt{SPIW} \cite{spiw}.

@@

%%%%%%%%% then the Bibliography, if any %%%%%%%%%
% \bibliographystyle{plain} % or "siam", or "alpha", etc.
\bibliographystyle{IEEEtran} % or "siam", or "alpha", etc.
% \nocite{*} % list all refs in database, cited or not
\bibliography{braker_bibs,Andersson_bibs,Yufans_bibs} % Bib database in "refs.bib"
